POMDP comparisons across sigma_m, sigma_g

### 2017-10-04 appendix/appendix.Rmd

```{r include=FALSE}
## aesthetics
knitr::opts_chunk$set(comment="", message=FALSE)
ggplot2::theme_set(ggplot2::theme_bw())
```

@@

We first define functions that correspond to each policy:

```{r}
msy_policy <- function(x) F_MSY * x
pgy_policy <- function(x) F_PGY * x

escapement_policy <- function(x){  ## ASSUMES harvest takes place *after* recruitment
  if(f(x,0) > B_MSY)
    f(x,0) - B_MSY
  else
    0
}
```

```{r}
# inits
x0 <- K/6
Tmax = 20

# simulate each policy, returning all results in a single unified data frame 
sims <- 
  list(msy = msy_policy, 
       det = escapement_policy, 
       pgy = pgy_policy) %>% 
  map_df(det_sim, 
         f, x0, Tmax, 
         .id = "method") 
```

```{r}
# plot simulation results 
sims %>%
  ggplot(aes(time, state, col=method)) + 
  geom_line() + 
  geom_hline(aes(yintercept=B_MSY), lwd=1) 
```

@@

@@

```{r}
states <- seq(0,2, length=100)
actions <- states

## Harvest after recruitment
det_policy <- sapply(states, function(x) pmax(f(x,0) - S_star,0))
det_action <- sapply(det_policy, function(x) which.min(abs(actions - x)))
msy_action <- sapply(states, function(x) which.min(abs(actions - x * F_MSY)))
pgy_action <- sapply(states, function(x) which.min(abs(actions - x * F_PGY)))

@@

```{r}
observations <- states # same discretization, for convenience

meta <- expand.grid(sigma_g = c(0.02, 0.1, 0.2), 
                    sigma_m = c(0, 0.1, 0.2),
                    stringsAsFactors = FALSE)

models <- 
  parallel::mclapply(1:dim(meta)[1], 
           function(i){
  fisheries_matrices(
  states = states,
  actions = actions,
  observed_states = observations,
  reward_fn = reward_fn,
  f = f,
  sigma_g = meta[i,"sigma_g"][[1]],
  sigma_m = meta[i,"sigma_m"][[1]],
  noise = "normal")
})

@@

@@

```{r}
# inits
x0 <- which.min(abs(states - K/6))
Tmax <- 100
reps <- 100
options(mc.cores = parallel::detectCores())

static_sims <-
  map_dfr(models, 
          function(m){
            do_policy_sim <- function(policy){
              sim <- sim_pomdp(m$transition, m$observation, m$reward, discount, 
                               x0 = x0, Tmax = Tmax, policy = policy, reps = reps)
              sim$df %>% 
                mutate(state = states[state], action = actions[action])
            }

            list(msy = msy_action, 
                 det = det_action, 
                 pgy = pgy_action) %>% 
              map_df(do_policy_sim, .id = "method")
          },
          .id = "scenario") 
```

```{r}
# add parameter metadata to the output data.frame to track which simulation had which parameters
meta_simple <-  
  meta %>% 
  select(sigma_m, sigma_g) %>%
  mutate(scenario  = as.character(1:dim(meta)[1]))
static_sims <- right_join(meta_simple, static_sims)
```

```{r}
# plot
static_sims %>%
  select(time, state, rep, method, sigma_m, sigma_g) %>%
  group_by(time, method, sigma_m, sigma_g) %>%
  summarise(mean = mean(state), sd = sd(state)) %>%
  ggplot(aes(time, mean, col=method, fill=method)) + 
  geom_line() + 
  geom_ribbon(aes(ymax = mean + sd, ymin = mean-sd), col = NA, alpha = 0.2) +
  facet_grid(sigma_m ~ sigma_g)
```

```{r eval=TRUE}
## log metadata associated with each run
log_data <- data.frame(id = id, 
                       model = "gs", 
                       r = r, 
                       K = K, 
                       sigma_g = sigma_g, 
                       sigma_m = sigma_m, 
                       noise = "normal")

dir.create(log_dir)

system.time({
alpha <- sarsop(m$transition, m$observation, m$reward, discount, 
                log_data = log_data, log_dir = log_dir,
                precision = .000002, timeout = 1000)
})
```

```{r}
meta <- meta_from_log(data.frame(id = id), log_dir)
alpha <- alphas_from_log(meta, log_dir)[[1]] ## bc fn returns a list with all matching alphas, we need [[1]]
```

## Combined plot

>>>>

```{r aethetics, include=FALSE}
knitr::opts_chunk$set(comment="", message=FALSE)
ggplot2::theme_set(ggplot2::theme_bw())
```

```{r helper_routines, include = FALSE}
det_simulations <- function(x0 = 1/6, Tmax = 20, 
                              B_MSY = K / 2, 
                              F_MSY = r / 2 , 
                              F_PGY = 0.8 * F_MSY){

  msy_policy <- function(x) F_MSY * x
  pgy_policy <- function(x) F_PGY * x
  escapement_policy <- function(x){  ## ASSUMES harvest takes place *after* recruitment
    if(f(x,0) > B_MSY)
      f(x,0) - B_MSY
    else
      0
  }

  # simulate each policy, returning all results in a single unified data frame 
  sims <- 
    list(msy = msy_policy, 
         det = escapement_policy, 
         pgy = pgy_policy) %>% 
    map_df(det_sim, 
           f, x0, Tmax, 
           .id = "method")

  sims
}
```

@@

```{r}
det_simulations(x0 = 1/6, Tmax = 20, B_MSY, F_MSY, F_PGY) %>%
  ggplot(aes(time, state, col=method)) + 
  geom_line() + 
  geom_hline(aes(yintercept=B_MSY), lwd=1) 
```

@@

```{r}
## Discretize space
states <- seq(0,2, length=100)
actions <- states
observations <- states
```

@@

```{r}
policies <- list(
  det = map_int(pmax(f(states,0) - S_star,0), ~ which.min(abs(actions - .x))),
  msy = map_int(states, ~ which.min(abs(actions - .x * F_MSY))),
  pgy = map_int(states, ~ which.min(abs(actions - .x * F_PGY))) )
```

@@

```{r}
meta <- expand.grid(sigma_g = c(0.02, 0.1, 0.2), 
                    sigma_m = c(0, 0.1, 0.2),
                    stringsAsFactors = FALSE) %>%
        mutate(scenario  = as.character(1:length(sigma_m)))
```

```{r}
models <- 
  parallel::mclapply(1:dim(meta)[1], 
           function(i){
  fisheries_matrices(
  states = states,
  actions = actions,
  observed_states = observations,
  reward_fn = reward_fn,
  f = f,
  sigma_g = meta[i,"sigma_g"][[1]],
  sigma_m = meta[i,"sigma_m"][[1]],
  noise = "normal")
})

@@

The POMDP solution is represented by a collection of alpha-vectors and values, returned in a `*.policyx` file.
Each scenario (parameter combination of `sigma_g`, `sigma_m`, and so forth) results in a separate solution file.

Because this solution is computationally somewhat intensive, we provide

```{r eval=FALSE}
log_dir <- "appendix_alphas" # Store the computed solution files here
dir.create(log_dir)
id = "pomdp_intro"    # an optional id for these runs

## POMDP solution (slow, >10,000 seconds per scenario, & memory intensive)
system.time(
  alphas <- 
    parallel::mclapply(1:length(models), 
    function(i){
      log_data <- data.frame(id = id, 
                             model = "gs", 
                             r = r, 
                             K = K, 
                             sigma_g = meta[i,"sigma_g"][[1]], 
                             sigma_m = meta[i,"sigma_m"][[1]], 
                             noise = "normal")

      sarsop(models[[i]]$transition,
             models[[i]]$observation,
             models[[i]]$reward,
             discount = discount,
             precision = 0.00000002,
             timeout = 10000,
             log_dir = log_dir,
             log_data = log_data)
    })
)

We can read the stored solution from the log:

```{r}
#alphas <- alphas_from_log(meta, log_dir)
```

@@

```{r static_sims}
Tmax <- 100
x0 <- which.min(abs(K/6 - states))
reps <- 100

static_sims <- 
 map_dfr(models, function(m){
            do_sim <- function(policy) sim_pomdp(
                        m$transition, m$observation, m$reward, discount, 
                        x0 = x0, Tmax = Tmax, policy = policy, reps = reps)$df
            map_dfr(policies, do_sim, .id = "method")
          }, .id = "scenario")

## Simulating the POMDP policies under uncertiainty

```{r pomdp_sims}
# slower!
pomdp_sims <- 
  map2_dfr(models, alphas, function(.x, .y){
             sim_pomdp(.x$transition, .x$observation, .x$reward, discount, 
                       unif_prior, x0 = x0, Tmax = Tmax, alpha = .y,
                       reps=reps)$df %>% 
              mutate(method = "pomdp") # include a column labeling method
           },
           .id = "scenario") %>% left_join(meta)
```

Combine the resulting data frames

```{r}
sims <- bind_rows(static_sims, pomdp_sims) %>%
    left_join(meta) %>%    ## combine output with scenarios
    mutate(state = states[state], action = actions[action]) # index -> value

write_csv(sims, "sims.csv")
```

##  Figure S1

We the results varying over different noise intensities, sigma_g, and sigma_m.  Figure 1 of the main text considers the case of sigma_g = 0.05, sigma_m = 0.1

```{r}
sims %>%
  select(time, state, rep, method, sigma_m, sigma_g) %>%
  group_by(time, method, sigma_m, sigma_g) %>%
  summarise(mean = mean(state), sd = sd(state)) %>%
  ggplot(aes(time, mean, col=method, fill=method)) + 
  geom_line() + 
  geom_ribbon(aes(ymax = mean + sd, ymin = mean-sd), col = NA, alpha = 0.2) +
  facet_grid(sigma_m ~ sigma_g)
```

### 2017-10-04 appendix/appendixB.Rmd
no printr

```{r libraries}
library(tidyverse)
library(nimble)
```

@@

```{r subset}
stock_ids <- c("PLAICNS", "ARGHAKENARG")
examples <- df2 %>% filter(stockid %in% stock_ids)
```

@@

  ## Return fit
  data.frame(stockid = fish$stockid[1],
             species = fish$commonname[1],
             r = mean(samples$r),
             K = mean(samples$K),
             sigma_g = mean(samples$sigma),
             r_sd = sd(samples$r),
             K_sd = sd(samples$K),
             sigma_g_sd = sd(samples$sigma),
             stringsAsFactors = FALSE)

@@

```{r}
pars <- fits %>% ungroup() %>% select(species, r, K, sigma_g) 
```

```{r}
# devtools::install_github("boettiger-lab/sarsop")  ## install package first if necessary.
library(sarsop)       # the main POMDP package
library(tidyverse)    # for munging and plotting
library(parallel)
options(mc.cores = 6) # Reserve ~ 10 GB per core
log_dir <- "pomdp_intro"
```

@@

```{r}
meta <- expand.grid(species = pars$species, 
                    sigma_m = c(0, 0.1, 0.2),
                    stringsAsFactors = FALSE) %>%
  left_join(pars)
meta
```

@@

```{r}
models <- 
  parallel::mclapply(1:dim(meta)[1], 
                     function(i){
                       fisheries_matrices(
                         states = states,
                         actions = actions,
                         observed_states = observations,
                         reward_fn = reward_fn,
                         f = gs(meta[i, "r"][[1]], meta[i, "K"][[1]]),
                         sigma_g = meta[i,"sigma_g"][[1]],
                         sigma_m = meta[i,"sigma_m"][[1]],
                         noise = "normal")
                     })

## log metadata associated with each run

dir.create(log_dir)

@@

```{r eval=FALSE}
## A generic routine to find stock size (x) which maximizes 
## growth rate (f(x,0) - x, where x_t+1 = f(x_t))

f <- gs(pars[i,"r"][[1]], pars[i, "K"][[1]])

S_star <- optimize(function(x) -f(x,0) + x / discount, 
                   c(0, 2*K))$minimum

B_MSY <- S_star
MSY <- f(B_MSY,0) - B_MSY

F_MSY <- MSY / B_MSY  
F_PGY = 0.8 * F_MSY

policies <- list(
  det = map_int(pmax(f(states,0) - S_star,0), ~ which.min(abs(actions - .x))),
  msy = map_int(states, ~ which.min(abs(actions - .x * F_MSY))),
  pgy = map_int(states, ~ which.min(abs(actions - .x * F_PGY))) )
```

>>>>

```{r libraries}
# devtools::install_github("boettiger-lab/sarsop")  ## install package first if necessary.
library(tidyverse)
library(sarsop)
library(nimble)
library(parallel)
```

@@

```{r subset}
stock_ids <- c("PLAICNS", "ARGHAKENARG")
examples <- df2 %>% 
  filter(stockid %in% stock_ids) %>% 
  ungroup() %>% 
  group_by(commonname)
```

@@

  ## Return fit
  data.frame(stockid = fish$stockid[1],
             commonname = fish$commonname[1],
             r = mean(samples$r),
             K = mean(samples$K),
             sigma_g = mean(samples$sigma),
             r_sd = sd(samples$r),
             K_sd = sd(samples$K),
             sigma_g_sd = sd(samples$sigma),
             stringsAsFactors = FALSE)

@@

```{r}
pars <- fits %>% ungroup() %>% select(commonname, r, K, sigma_g) 
```

```{r}

options(mc.cores = 6) # Reserve ~ 10 GB per core
log_dir <- "pomdp_intro"
```

@@

```{r}
meta <- expand.grid(commonname = pars$commonname, 
                    sigma_m = c(0, 0.1, 0.2),
                    stringsAsFactors = FALSE) %>%
  left_join(pars) %>%
  mutate(scenario  = as.character(1:length(sigma_m)))

meta
```

@@

```{r}
models <- 
  parallel::mclapply(1:dim(meta)[1], 
                     function(i){
                       fisheries_matrices(
                         states = states,
                         actions = actions,
                         observed_states = observations,
                         reward_fn = reward_fn,
                         f = gs(meta[i, "r"][[1]], meta[i, "K"][[1]]),
                         sigma_g = meta[i,"sigma_g"][[1]],
                         sigma_m = meta[i,"sigma_m"][[1]],
                         noise = "normal")
                     })
```

```{r}
dir.create(log_dir)

@@

```{r}
pars <- examples %>% 
  group_by(commonname) %>% 
  summarise(N = max(SSB)) %>% 
  right_join(pars)
```

<!--- ICK, find a more streamlined way to do this stuff species by species -->

Convert example data into discrete index space.

```{r}
ex <- examples %>%
  mutate(biomass = index(scaled_biomass, states),          
         catch = index(scaled_catch, actions))  %>% 
  select(year, biomass, catch, commonname)
```

Add corresponding static policy levels on:

```{r}
statics <- function(P){
  f <- gs(P$r, P$K)
  S_star <- optimize(function(x) -f(x,0) + x / discount, c(0, 2* P$K))$minimum
  B_MSY <- S_star
  MSY <- f(B_MSY,0) - B_MSY

  tibble(S_star, F_MSY = MSY / B_MSY, F_PGY = 0.8 * F_MSY, 
         commonname = P$commonname, N = P$N)
  }

  pars %>% 
  rowwise() %>% 
  do(statics(.)) %>% 
  right_join(ex) %>% 
  ungroup()  -> ex2

  ex2
```

```{r} 
det_f <- function(S_star, i) 
  map_int(pmax(f(states,0) - S_star[[1]],0), ~ which.min(abs(actions - .x)))[i]
msy_f <- function(F_MSY, i) 
  map_int(states, ~ which.min(abs(actions - .x * F_MSY[[1]])))[i]
pgy_f <- function(F_PGY, i) 
  map_int(states, ~ which.min(abs(actions - .x * F_PGY[[1]])))[i]

rescale <- function(x, N) states[x]*N

historical <- ex2 %>% 
  group_by(commonname) %>% 
  mutate(det =  det_f(S_star, biomass),
         msy =  msy_f(F_MSY, biomass),
         pgy =  pgy_f(F_PGY, biomass)) %>%
  mutate_if(is.integer, rescale, N=N) %>%
  select(year, biomass, catch, det, msy, pgy, commonname)
```

```{r}
historical %>%
select(year, commonname, pgy, msy, det) %>%
gather(model, stock, -year, -commonname) %>%
  ggplot(aes(year, stock, col=model)) + 
  geom_line(data = historical, aes(year, biomass), col="grey", lwd=1) +
  geom_line(data = historical, aes(year, catch), col="black", lwd=1) +
  geom_line(lwd=1) + facet_wrap(~commonname, scales = "free", ncol=1) + theme_bw()  
```

## POMDP historical

```{r pomdp_sims}
## FIXME obs, action is in index space!
pomdp_sims <- 
  map2_dfr(models, alphas, function(.x, .y){
             hindcast_pomdp(.x$transition, .x$observation, .x$reward, discount, 
                            obs = historical$scaled_biomass, actions = historical$scaled_catch,
                            state_prior = unif_prior,
                            alpha = .y)$df %>% 
              mutate(method = "pomdp") # include a column labeling method
           },
           .id = "scenario") %>% left_join(meta)
```
